\textbf{International Transport costs:\\New Findings from modeling additive costs}\\Online Appendix (Not for Publication)

### 2021-11-15 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
début relecture

\begin{abstract}
International transport costs do have an additive part. By how much? To what extent it matters? This paper provides new answers to these questions. Using information contained in the US imports flows from 1974 to 2019, we develop an empirical model that disentangles the ad-valorem and the additive components of international transport costs. The per-unit component of transport costs represents a sizeable share of total transport costs, between 30\% and 45\% depending on the year and the transport mode considered. We then investigate the important consequences of modelling additive costs, under two different perspectives. First, modelling varying additive costs modifies the decomposition of transport costs time trend between the reduction in ``pure'' transport costs and trade composition effects, the latter playing a minor role. Second, we revisit the welfare gains of trade liberalisation in presence of additive costs. In this regard, we shed light on the welfare variations induced by the ``hyper-globalization'' period (and the subsequent retrenchment after 2008-2009), as well as the disproportionate part of additive transport costs in determining those welfare variations. Neglecting the additive component substantially underestimates the welfare gains of the transport cost decrease, by around 1.2 percentage points. %In both aspects, our results point to the importance of modelling the additive component of international transport costs.
\end{abstract}

@@

Trade costs are central to international economic analysis. In particular, they are known to be a major obstacle to international economic integration and international trade flows. Defined as the costs associated with the exchange of goods across
national borders, trade costs are usually split between transaction costs, policy costs, time costs and transport costs \textit{per se}. Many argue that the latter have substantially decreased with technological advance in transportation, infrastructure development
and new communication technologies. However, several empirical evidence
suggest that transport costs remain large and deserve attention.\footnote{\cite{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs. Equivalently, international transportation costs represent a 21\% markup over production costs. Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows. \cite{Disdier_Head08}, among others, point out that distance as an incompressible cost is still a very important determinant of trade flows.} In addition, progress in transportation techniques (which reduces the constant-dollar cost of transport) should be compared with technical progress in production (which reduces the constant-dollar cost of production) to get a clear picture of the evolution
of the transport costs burden. The latter also depends on whether the cost is in proportion of the value of goods traded (hence does not distort relative prices) or a cost per unit traded (the additive form conversely affecting relative prices). If a complex task, having a clear view of the international transport cost burden is of key importance, since the answers to several important (both academic and policy) economic issues depend in large part on how much distance-based costs affects trade, such as, among others, the dynamics of agglomeration, the profitability of outsourcing or the gains from trade cost reduction. Our main contribution in this paper is to bring new insights on this matter.\smallskip

@@

To what extent should we care about additive costs? Our paper answers this question from two different perspectives.
On the empirical side, we show the importance of incorporating additive costs for properly analyzing the decreasing trend in international transport costs observed between 1974 and 2019. By allowing for time-varying additive costs, we find that the downward trend in international transport costs comes from the reduction in ``pure'' transport costs (i.e., for a given product/country partner) over time, by roughly 60\% in both air and vessel shipping over 1974-2019. Trade composition effects (i.e., changes over time in the baskets of imported goods and/or the origin countries) have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs, in contrast with the conclusion obtained by \cite{hummels2007} yet also on US import flows.
%\cite{hummels2007} finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs, because trade composition effects have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs.
Importantly, we show that the difference of results can be attributed to the new method of modeling the additive component that we offer. Allowing the share of additive costs to vary across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.\smallskip

>>>>

\begin{abstract}
International transport costs do have an additive part. By how much? To what extent it matters? This paper provides new answers to these questions. Using information contained in the US imports flows from 1974 to 2019, we develop an empirical model that disentangles the ad-valorem and the additive components of international transport costs. The per-unit component of transport costs represents a sizeable share of total transport costs, between 30\% and 45\% depending on the year and the transport mode considered. We then investigate the important consequences of modelling additive costs, under two different perspectives. First, modelling varying additive costs modifies the decomposition of transport costs time trend between the reduction in ``pure'' transport costs and trade composition effects, the latter playing a minor role. Second, we revisit the welfare gains of trade liberalisation in presence of additive costs. In this regard, we shed light on the welfare variations induced by the ``hyper-globalization'' period, as well as the disproportionate part of additive transport costs in determining those welfare variations. Neglecting the additive component substantially underestimates the welfare gains of the transport cost decrease, by around 1.2 percentage points. 
\end{abstract}

@@

Trade costs are central to international economic analysis. In particular, they are known to be a major obstacle to international economic integration and international trade flows. Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs, policy costs, time costs and transport costs \textit{per se}. Many argue that the latter have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies. 
However, several empirical evidence suggest that transport costs remain large and deserve attention.\footnote{\cite{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs. Equivalently, international transportation costs represent a 21\% markup over production costs. Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) attests to the impact of transport costs on trade flows. \cite{Disdier_Head08}, among others, point out that distance as an incompressible cost is still a very important determinant of trade flows.} In addition, progress in transportation techniques (which reduces the constant-dollar cost of transport) should be compared with technical progress in production (which reduces the constant-dollar cost of production) to get a clear picture of the evolution
of the transport costs burden. The latter also depends on whether the cost is in proportion of the value of goods traded (hence does not distort relative prices) or a cost per unit traded (the additive form conversely affecting relative prices). If a complex task, having a clear view of the international transport cost burden is of key importance, since the answers to several important (both academic and policy) economic issues depend in large part on how much distance-based costs affects trade, such as, among others, the dynamics of agglomeration, the profitability of outsourcing or the gains from trade cost reduction. Among others, an important contribution in this paper is to bring new insights on the latter issue.\smallskip

@@

To what extent should we care about additive costs? Our paper answers this question from two different perspectives.
On the empirical side, we show the importance of incorporating additive costs for properly analyzing the decreasing trend in international transport costs observed between 1974 and 2019. By allowing for time-varying additive costs, we find that the downward trend in international transport costs comes from the reduction in ``pure'' transport costs (i.e., for a given product/country partner) over time, by roughly 60\% in both air and vessel shipping over 1974-2019. Trade composition effects (i.e., changes over time in the baskets of imported goods and/or the origin countries) have only limited influence in accounting for the downward trend of international transport costs, and when they matter (in maritime transport), composition effects amplify the reduction of pure transport costs, in contrast with the conclusion obtained by \cite{hummels2007} also on US import flows.
%\cite{hummels2007} finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs, because trade composition effects have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs.
Importantly, we show that the difference of results can be attributed to the new method of modeling the additive component that we offer. Allowing the share of additive costs to vary across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period.\smallskip
